Numerical simulation of bubble deformation \\ and breakup under simple linear shear flows

### 2024-11-29 Revised-paper/Reply/Reply-Reviewer1.tex

Please find enclosed the revised version of our manuscript (CES-D-24-01867):
``Numerical Simulation of Bubble Deformation and Breakup under 
Simple Linear Shear Flows''.
We thank the reviewers for their comments and thoughts regarding improvement 
of our paper. We believe that we have addressed all of the reviewers’ 
concerns; the changes are itemized in detail below.

\par\noindent
NOTE: all references to equations and figures below are with
reference to the numbering scheme of the revised version of the paper,
not the original version.  Also, changes are hi-lighted in red.
\par\noindent

Changes made in response to comments of Reviewer 1: \begin{enumerate}
%%1
\item
\textsf
{Abstract, hightlights, and introduction are written in a way that confuses the reader: the text jumps back and forth between bubbles and drops as well as shear flows and other flow configurations. Then it suddenly moves to lift and deformation rather than breakup. Finally a number of random applications are mentioned without any relation to bubble breakup. The authors should clearly state what their contrbution is from the outset. The reader will not go on until the end to figure that out by himself. The use of shorter paragraphs each addressing a single issue is very helpful.}
\vspace{3 mm}

Response: \\
We have gone through the whole paper fixing the grammar and structure.  We believe the message is very clear now. \\
\\

@@

%%2
\item
\textsf
{Method: The use of only 2 levels of grid refinement appears odd. For an accurate representation of the flow near the interface a very fine mesh is typically needed. Only 32 cells over the bubble diameter is hardly sufficient ot get accurate results as corroborated by the mediorce agreement of the presented simulation results with the experimental data used for validation. Two more levels of grid refinement are likely needed to resolved the flow near the bubble while farther away some coarsening may be possible. Optimizing the grid is imperative in a situaltion like the considerd one due to the strong deformation of the bubble which high curvature of the interface and the singular nature of the breakup event. Fully capturing the breakup, which represents a topological singularity, is impossible without a subgrid scale model.}
\vspace{3 mm}

Response: \\
We totally disagree with the referee here.  The resolution used for our results is enough.  The results shown in Figure 3 are very convincing in that if we refined the grid more, the results would hardly change.  In addition, we have the results in Tables I and II which compare our results to other researchers' simulations and experiments for bubble/drop deformation in shear flows. Finally, at the top of section 3.2, we now list a number of references involving bubbles and drops with severe deformation in which our method was validated by way of grid refinement studies and comparison with experiments.

There is a big gaping hole in the research in which our results fit in: this is the low Capillary number, Moderate to High Reynolds number regime.  In this void, it is impossible to do experiments unless one is in a microgravity environment.  At the same time, numerical algorithms must be designed in order to robustly and accurately predict the tensile strength of a bubble in the low capillary, moderate to high Reynolds number regime, in which ours is the first.  We believe that the results as currently presented, fill the void and that they should be reported now rather than waiting another year.

Perhaps in future work, in which we have more computing power available to us, we can try an extra refinement, albeit, the results are most likely a foregone conclusion in that they will be in close agreement with the results that we already have.  Also as future work, it is inevitable that researchers will be able to run experiments on the space station which corroborate the results of our simulations that we have now.

We have also added ``Remark 1'' and ``Remark 2'' in section 3.1 which makes it clear why adding another level of refinement would delay the publication of our work by another four years, and why sub-scale models are unnnecessary at pinch off for the bubble deformation under simple shear flow problem.  Again, the results that we have now are publishable and will make an impact, it would be fruitless to wait another 4 years.

@@

%%3
\par\noindent
\item
\textsf
{Results: Mostly just pictures are shown from which little is to be learned. The only results of which future use can be expected is the regime map in Fig. 12. Bu this is also only qualitative and based on very few data points. More combinations of Ca and Re should be provided and a quantitative fit to the regime boundary be obtained. Moreover only two values of density and viscosity ratio equal to 1 and ~0.001 are used. An investigation of intermediate values would be straight forward and would allow to provide also models for the influence of these obviously important parameters. In particular it should be clarified whether the difference in behavior is more due to the density ration or more due to the viscosity ratio. Also it needs to be investigated how small these ratios need to be for the gas phase properties to be considered as negligibly small. For the practically important case of air bubbles in water the viscosity ratio is 0.02, an order of magnitude larger than considered in the paper. Whether this can already be considered negligible or not remains unclear.}
\vspace{3 mm}

Response: \\
We totally disagree with this comment too.  We already report many novel and significant results and sufficiently analyze the data both with strategic plots and quantitative tables.  Adding more would only needlessly delay the paper another year.  
\par
We do agree with the referee in that the recommended new simulations would be an excellent inclusion in future papers.  Please see our modified conclusion, last paragraph.

@@

%%4
\par\noindent
\item
\textsf
{The effect of gravity is neglected which potentially obliterates the results. It is not at all clear that the effect of the shear flow can be separated from that of gravity without rendering the results meaningless for any practical application. Even when one wants to do so it needs to be considered that the bubbles will be deformed into an ellipsoidal shape. Then at least the influence of initial condition has to be checked in order to ensure that the results make any sense. \\}
\vspace{3 mm}

Response: \\
We totally disagree that not including gravity ``obliterates'' the results.

At the least, there are a host of strictly microgravity applications in which our research will be of great help.   We list many new applications, including the microgravity applications, in the revised version: please see the first paragraph in the introduction.  All these applications that we list will significantly benefit from our article.

We have also added a discussion at the end of section 2 (see the ``Remark'') in which we make some predictions as to how the inclusion of gravity would change the results.

Having said that, we do agree that taking into account the gravity effects and initial condition effects will make good material for future work.  Please see the last paragraph in our modified conclusion.

\\

@@

%%5
\par\noindent
\item
\textsf
{Language and writing (examples):\\
line 191 and following: "Timestep $\Delta t$" appears out of any context. So do "Step 1." to "Step 4".\\
line 232: ""bubble deformation in simple linear shear flow" results" is impossible language. \\}
\vspace{3 mm}

Response: \\
We have fixed these items.  Also, we have made many complete ``run-throughs'' of our paper in order to remove all mistakes and awkward sentences.

### 2024-11-29 Revised-paper/Paper/Bubble-deformation-Rev.tex
replies

\begin{abstract}
\textcolor{red}
{
Numerical simulations are presented for the deformation and breakup of a bubble in liquids undergoing simple linear shear flow in a variety of regimes including the low Capillary number, high Reynolds number regime.  Numerical results are obtained using a projection method for incompressible two-phase flow. The method represents interfaces using the sharp interface coupled level set and Volume-Of-Fluid (CLSVOF) method.   To verify the CLSVOF numerical algorithm and provide a basis for comparison, computational results are also presented that compare with previous, more popular, drop deformation and rupture simulations.  The numerical results reveal that the shear-induced bubble deformation and breakup process differs significantly from that of the analogous drop system under similar flow conditions.  The first distinguishing factor for bubble breakup is that the magnitude of shear flow necessary for causing rupture is much larger than that in the drop system case.  In other words, a larger Reynolds number is needed to induce bubble breakup.  The second distinct feature of the bubble system versus the drop system is that the bubble does not maintain a stable deformed shape as the system parameters approach the critical ``breakup'' Reynolds number.  It is asserted that the differences in morphology for a bubble undergoing breakup versus a drop in the same process can be attributed to the density and viscosity ratio of the corresponding two-phase flow systems.  The critical conditions for bubble breakup with respect to the Reynolds and capillary numbers are determined for several cases. }
\end{abstract}

@@

	This article presents computational studies of shear-driven deformation and breakup of a bubble in insoluble viscous liquids.  Studying bubble break-up, in which we focus only on the balance of the wall driving force and bubble surface tension force, via computation rather than experiments simplifies the process of setting a combination of precise, simple shear flow conditions, low Capillary number ($Ca=\mu_{m}U/\sigma$) conditions, high Reynolds' number conditions ($Re=\rho_{m} R U/\mu_{m}$), low-density ratio, low viscosity ratios, and zero buoyancy effects.  We remark that previous controlled experimental studies on shear-driven bubble deformation are restricted to the low Reynolds' number and large capillary number regime\cite{RusMan02,CanedoETAL}.  One reason is that buoyancy effects are minimized in highly viscous liquids.  On the other hand, many applications are characterized by liquids in the high Reynolds' number regime.  The physical properties that distinguish bubble and drop studies are expressed in terms of the density ratio $\lambda = \rho_{\rm b} / \rho_{\rm m}$ and the viscosity ratio $\eta = \mu_{\rm b} / \mu_{\rm m}$, where $\rho$ is the fluid density, $\mu$ is the viscosity and the subscripts ``b'' and ``m'' denote the ``bubble'' or ``drop'' and the ``matrix fluid'', respectively.  For a bubble in an insoluble, viscous liquid, $\lambda \simeq 0$ and $\eta \simeq 0$, whereas most studies dealing with a drop in an immiscible viscous liquid take $\lambda =1$ and $\eta \simeq 1$.  

@@

{ Thus, a lot of studies about the deformation and breakup of a drop in simple linear shear have been presented so far. In contrast, few studies have been conducted on bubble deformation and breakup.  In the low Capillary number, high Reynolds' number regime, there are no controlled experiments or simulations.  We reiterate why there have been few studies regarding the ``tensile strength'' of bubbles. Experimentally, if one wants to isolate the interplay of shearing force with the bubble surface tension force, in the moderate to high Reynolds number regime, and low Capillary number regime, one is restricted to microgravity conditions.  Computational experiments are difficult too. In order to accurately compute the tensile strength of a bubble, one must resort to a combination of parallel computing, the multigrid preconditioned conjugate gradient method\cite{tatebe1993multigrid,SusSmiHusOhtZhi07} for poorly conditioned large sparse matrix systems, adaptive mesh refinement\cite{AMReX_JOSS,SusSmiHusOhtZhi07}, and a robust, volume preserving interface tracking method (we use the CLSVOF method\cite{SusPuc00,SusSmiHusOhtZhi07}).  These aspects of our algorithm allow one to accurately simulate the balance of the restoring bubble surface tension force against the driving wall shear force and at the same time, accurately simulate the evolution of the bubble shape to bubble-break-up if the given shear stress and physical properties permit it.

@@

\subsection{Validation of the numerical method}
%  -----------------------------------------------------------------------------
\textcolor{red}
{
	The effectiveness of our sharp interface computational method has been demonstrated via grid refinement studies and comparison with experiments for the complicated rising motion of single bubbles and drops in viscous liquids \citet{OhtSus12, OhtAkaYosSus14, OhtFurYosSus19,ohta2010sensitivity,stewart2008improved,SusSmiHusOhtZhi07}, the simulation of atomization in a realistic diesel injector\cite{arienti2013coupled}, and the simulation of bubble formation due to the injection of gas through a nozzle\cite{ohta2011robust}.  In this section, the accuracy of our computational method will be verified for the problem of shear-induced deformation of a drop and bubble. 
}
\par
       First, we compare quantitatively against the steady-state drop deformation results reported by \citet{LiRenRen00}.  The shape of a deformed drop in simple linear shear flow is described in terms of the Taylor deformation parameter $D$=$(a-b)/(a+b)$, where $a$ and $b$ are the major and minor axes of the deformed drop, respectively.  For consistency, we perform numerical simulations using CLSVOF over the same computational domain and grid size used in \citet{LiRenRen00}, which has dimensions $\lwh{8}{4}{8}$ (recall that $R$ is the bubble/drop radius) and a level-0 grid size $\Delta x=\Delta y=\Delta z=R/8$; our two-level AMR grid structure also uses a finer level-1 grid size $\Delta x^{\ell=1} = \Delta y^{\ell=1} = \Delta z^{\ell=1} = R/16$.  Numerical results are listed in Table~\ref{tab:DeComparison} for $D$ as a function of $Re$, with $Ca=0.3$ and $\lambda = \eta = 1$ fixed in every case.  The results in Table~\ref{tab:DeComparison} compare computations using our CLSVOF algorithm with corresponding results obtained with the VOF method used in \citet{LiRenRen00}.  
%
\begin{table}[tbh]
\caption{Comparison of the Taylor deformation parameter $D$ for a drop as a function of $Re$ ($Ca=0.3$, $\lambda = \eta = 1$). In all cases, the domain size is $\lwh{8}{4}{8}$, where $R$ is the drop radius.  The CLSVOF method computations used a two-level AMR computational domain with a level-0 discretization $\Delta x^{\ell=0} = \Delta y^{\ell=0} = \Delta z^{\ell=0} = R/8$ and a level-1 discretization $\Delta x^{\ell=1} = \Delta y^{\ell=1} = \Delta z^{\ell=1} = R/16$.}
\label{tab:DeComparison}
%\footnotesize
\center
\begin{tabular}{ c  c  c  c  c }
\hline
\hline
Reynolds number                      & 0.1     & 0.5     & 0.6     & 0.75      \\
$D$(\citet{LiRenRen00})  & 0.3968  & 0.45    & 0.4768  & Breakup   \\
$D$(Our study) & 0.3960  & 0.4570  & 0.4758  & Breakup   \\
\hline
\hline
\end{tabular}
\end{table}

>>>>

\begin{abstract}
\textcolor{red}
{
Numerical simulations are presented for the deformation and breakup of a bubble in liquids undergoing simple linear shear flow in a variety of regimes including the low Capillary number, moderate/high Reynolds number regime.  Numerical results are obtained using a projection method for incompressible two-phase flow. The method represents interfaces using the sharp interface coupled level set and Volume-Of-Fluid (CLSVOF) method.   To verify the CLSVOF numerical algorithm and provide a basis for comparison, computational results are also presented that compare with previous, more popular, drop deformation and rupture simulations.  The numerical results reveal that the shear-induced bubble deformation and breakup process differs significantly from that of the analogous drop system under similar flow conditions.  The first distinguishing factor for bubble breakup is that the magnitude of shear flow necessary for causing rupture is much larger than that in the drop system case.  In other words, a larger Reynolds number is needed to induce bubble breakup.  The second distinct feature of the bubble system versus the drop system is that the bubble does not maintain a stable deformed shape as the system parameters approach the critical ``breakup'' Reynolds number.  It is asserted that the differences in morphology for a bubble undergoing breakup versus a drop in the same process can be attributed to the density and viscosity ratio of the corresponding two-phase flow systems.  The critical conditions for bubble breakup with respect to the Reynolds and capillary numbers are determined for several cases. }
\end{abstract}

@@

	At the present, there is a large hole in the research concerning fundamental bubble dynamics in a simple shear flow.  There are no results, either experimentally or numerically, in the low Capillary number ($Ca=\mu_{m}U/\sigma$), moderate/high Reynolds number ($Re=\rho_{m} R U/\mu_{m}$) regimes.  In this void, it is impossible to do experiments unless one is in a microgravity environment.  At the same time, numerical algorithms must be designed in order to robustly and accurately predict the tensile strength of a bubble in the low capillary, moderate to high Reynolds number regime, in which ours is the first to be reliably used for such a purpose. 
	This article presents computational studies of shear-driven deformation and breakup of a bubble in insoluble viscous liquids.  Studying bubble break-up, in which we focus only on the balance of the wall driving force and bubble surface tension force, via computation rather than experiments simplifies the process of setting a combination of precise, simple shear flow conditions, low Capillary number ($Ca=\mu_{m}U/\sigma$) conditions, moderate/high Reynolds' number conditions ($Re=\rho_{m} R U/\mu_{m}$), low-density ratio, low viscosity ratios, and zero buoyancy effects.  We remark that previous controlled experimental studies on shear-driven bubble deformation are restricted to the low Reynolds' number and large capillary number regime\cite{RusMan02,CanedoETAL}.  One reason is that buoyancy effects are minimized in highly viscous liquids.  On the other hand, many applications are characterized by liquids in the moderate/high Reynolds' number regime.  The physical properties that distinguish bubble and drop studies are expressed in terms of the density ratio $\lambda = \rho_{\rm b} / \rho_{\rm m}$ and the viscosity ratio $\eta = \mu_{\rm b} / \mu_{\rm m}$, where $\rho$ is the fluid density, $\mu$ is the viscosity and the subscripts ``b'' and ``m'' denote the ``bubble'' or ``drop'' and the ``matrix fluid'', respectively.  For a bubble in an insoluble, viscous liquid, $\lambda \simeq 0$ and $\eta \simeq 0$, whereas most studies dealing with a drop in an immiscible viscous liquid take $\lambda =1$ and $\eta \simeq 1$.  

@@

{ Thus, a lot of studies about the deformation and breakup of a drop in simple linear shear have been presented so far. In contrast, few studies have been conducted on bubble deformation and breakup.  In the low Capillary number, moderate/high Reynolds' number regime, there are no controlled experiments or simulations.  We reiterate why there have been few studies regarding the ``tensile strength'' of bubbles. Experimentally, if one wants to isolate the interplay of shearing force with the bubble surface tension force, in the moderate to high Reynolds number regime, and low Capillary number regime, one is restricted to microgravity conditions.  Computational experiments are difficult too. In order to accurately compute the tensile strength of a bubble, one must resort to a combination of parallel computing, the multigrid preconditioned conjugate gradient method\cite{tatebe1993multigrid,SusSmiHusOhtZhi07} for poorly conditioned large sparse matrix systems, adaptive mesh refinement\cite{AMReX_JOSS,SusSmiHusOhtZhi07}, and a robust, volume preserving interface tracking method (we use the CLSVOF method\cite{SusPuc00,SusSmiHusOhtZhi07}).  These aspects of our algorithm allow one to accurately simulate the balance of the restoring bubble surface tension force against the driving wall shear force and at the same time, accurately simulate the evolution of the bubble shape to bubble-break-up if the given shear stress and physical properties permit it.

@@

\subsection{Validation of the numerical method}
%  -----------------------------------------------------------------------------
\textcolor{red}
{
	The effectiveness of our sharp interface computational method has been demonstrated via grid refinement studies and comparison with experiments for the complicated rising motion of single bubbles and drops in viscous liquids \citet{OhtSus12, OhtAkaYosSus14, OhtFurYosSus19,ohta2010sensitivity,stewart2008improved,SusSmiHusOhtZhi07,sussman2009stable}, the simulation of atomization in a realistic diesel injector\cite{arienti2013coupled}, and the simulation of bubble formation due to the injection of gas through a nozzle\cite{ohta2011robust}.  In this section, the accuracy of our computational method will be verified for the problem of shear-induced deformation of a drop and bubble. 
}
\par
       First, we compare quantitatively against the steady-state drop deformation results reported by \citet{LiRenRen00}.  The shape of a deformed drop in simple linear shear flow is described in terms of the Taylor deformation parameter $D$=$(a-b)/(a+b)$, where $a$ and $b$ are the major and minor axes of the deformed drop, respectively.  For consistency, we perform numerical simulations using CLSVOF over the same computational domain and grid size used in \citet{LiRenRen00}, which has dimensions $\lwh{8}{4}{8}$ (recall that $R$ is the bubble/drop radius) and a level-0 grid size $\Delta x=\Delta y=\Delta z=R/8$; our two-level AMR grid structure also uses a finer level-1 grid size $\Delta x^{\ell=1} = \Delta y^{\ell=1} = \Delta z^{\ell=1} = R/16$.  Numerical results are listed in Table~\ref{tab:DeComparison} for $D$ as a function of $Re$, with $Ca=0.3$ and $\lambda = \eta = 1$ fixed in every case.  The results in Table~\ref{tab:DeComparison} compare computations using our CLSVOF algorithm with corresponding results obtained with the VOF method used in \citet{LiRenRen00}.  
%
\begin{table}[tbh]
\caption{Comparison of the Taylor deformation parameter $D$ for a drop as a function of $Re$ ($Ca=0.3$, $\lambda = \eta = 1$). In all cases, the domain size is $\lwh{8}{4}{8}$, where $R$ is the drop radius.  The CLSVOF method computations used a two-level AMR computational domain with a level-0 discretization $\Delta x^{\ell=0} = \Delta y^{\ell=0} = \Delta z^{\ell=0} = R/8$ and a level-1 discretization $\Delta x^{\ell=1} = \Delta y^{\ell=1} = \Delta z^{\ell=1} = R/16$.}
\label{tab:DeComparison}
%\footnotesize
\center
\begin{tabular}{ c  c  c  c  c }
\hline
\hline
Reynolds number                      & 0.1     & 0.5     & 0.6     & 0.75      \\
$D$(\citet{LiRenRen00})  & 0.3968  & 0.45    & 0.4768  & Breakup   \\
$D$(Our study) & 0.3960  & 0.4570  & 0.4758  & Breakup   \\
\hline
\hline
\end{tabular}
\end{table}
